Generate Real and Pseudo-absences

### 2024-01-09 scripts/generate_absences.Rmd
updated lda dropper to deal with 0 variance variables

out1 <- lapply(records, absence_drawer, terrestrial = land,  bg_abs = 0.15)

out <- bind_rows(out1) %>% 
  rename(taxon = Taxon) %>% 
  mutate(date = as.Date('2023-12-18'), .before = geometry)

st_write(out, '../data/raw/occurrence/absences/absences.shp', append = FALSE)

@@

```{r import raster stack}

p <- '/media/steppe/hdd/Geospatial_data/WesternPlantPredictors2/chelsa'
paths <- file.path(p, list.files(p, recursive = F))
files <- lapply(paths, function(x){ file.path(x, list.files(x)) })
names(files) <- basename(paths)
bio <- lapply(files, vrt)
bio <- rast(bio)
bio <- subset(bio, c(1:11, 14:17, 19)) # these 'empty' need investigate

@@

write_rds(collinear_probs, '../results/bioclim_VIF_PA_drop.rda')

@@

collinear_probs <- read_rds('../results/bioclim_VIF_PA_drop.rda')
bio <- usdm::exclude(bio, collinear_probs)

@@

>>>>

out1 <- lapply(records, absence_drawer, terrestrial = land,  bg_abs = 0.15, rand_dist = 1000)

out <- bind_rows(out1) %>% 
  rename(taxon = Taxon) %>% 
  mutate(date = as.Date('2023-01-09'), .before = geometry)

st_write(out, '../data/raw/occurrence/absences/absences1km.shp', append = FALSE)

@@

```{r import raster stack, eval = F}

p <- '~/Documents/WesternPlantPredictors2'
paths <- file.path(p, list.files(p, recursive = F))
files <- lapply(paths, function(x){ file.path(x, list.files(x)) })
names(files) <- basename(paths)
bio <- lapply(files, vrt)
bio <- rast(bio)
rm(p, paths, files)

```

```{r read in predictors2 }
p <- '~/Documents/WesternPlantPredictors2'
paths <- list.dirs(p, recursive = T)
paths <- paths[2:length(paths)]
paths <- paths[! grepl('chelsa$|veg_cover$|soilgrids$|sand$|silt$|clay$|phh2o$|cfvo$|soc$', paths) ]

files <- lapply(paths, function(x){ file.path(x, list.files(x)) })
names(files) <- basename(paths)

layers <- lapply(files, vrt)
layers <- rast(layers)
names(layers) <- basename(paths)

rm(paths, files)
```

write_rds(collinear_probs, '../results/VIF_PA_drop.rda')

@@

```{r subset rasterstack 2, eval = F}
vif_vals <- spatSample(layers, size = 5000, values = TRUE) # this steps takes forever... 
vif_vals <- vif_vals[ complete.cases(vif_vals), ] # 4148 records used
collinear_probs <- usdm::vifstep(vif_vals, th = 10)
layers <- usdm::exclude(layers, collinear_probs)

write_rds(collinear_probs, '../results/VIF_PA_drop.rda')

@@

collinear_probs <- read_rds('../results/VIF_PA_drop.rda')
bio <- usdm::exclude(bio, collinear_probs)

@@

set.seed(27)
cn <- names(layers)
splicies <- split(recs, f = recs$taxon)

lapply(splicies[41:335], lda_PA_dropper, path = '../results/lda2/', col_names = cn)

which(names(splicies) == 'Asclepias speciosa')
